---
title: "assignment 1 question 2"
format: html
---

### 2. Proofs & SVD Example
#### A. Prove that $A \times B \neq B \times A$
```{r commutative}
# --------------
# AB <> BA
# --------------

# define two 3x3 matrices A and B

A <- matrix(c(1,2,3,
              4,5,6,
              7,8,9),
            nrow = 3, byrow = TRUE)

B <- matrix(c(1, 0,-1,
              1, 2, 0,
              0,-1, 1),
            nrow = 3, byrow = TRUE)

A%*%B
B%*%A

```


#### B. Prove that $A^T*A$ is always symmetric.

The matrix $A^T \times A$ is always symmetric because by transposing any matrix $A$ with dimensions $m \times n$, we get a matrix $A^T$ with dimensions $n \times m$. This has two implications for multiplying $A^T \times A$:

1. The multiplication $(n \times m) \times (m \times n)$ always works because the number of columns in the first matrix ($m$) match the number of rows($m$) in the second, which makes matrix multiplication possible.

2. The resulting matrix has dimensions $n \times n$, which is square and symmetric, because the dimensions are defined by the number of rows in the first matrix ($n$) and the number of columns in the second matrix (also $n$).

```{r symmetric}

# Matrix A from above:
A

# Transpose of A:
t(A)

# Transpose of A times A:
t(A) %*% A

```



#### C. Prove that $det(A^T \times A)$ is non-negative
$A^T \times A$ results in a square matrix as shown above, so we can calculate a determinant. 

And based on the theorem that $det(A^T \times A) = det(A)^2$, it follows that the determinant of $A^T \times A$ cannot be negative because any number times itself must be $\geq 0$.

Therefore, if  $det(A^T \times A) = det(A)^2$ and $det(A)^2 \geq 0$ then $det(A^T \times A) \geq 0$.

```{r determinant}

# determinant of transpose of A times A:
det(t(A) %*% A)

```


#### D. Singular Value Decomposition (SVD) and Image Compression: Write an R function that performs Singular Value Decomposition (SVD) on a grayscale image (which can be represented as a matrix). Use this decomposition to compress the image by keeping only the top k singular values and their corresponding vectors. Demonstrate the effect of different values of k on the compressed image's quality.
